---
title: "Untitled"
author: "Shasta Webb"
date: "2024-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### LOADING REQUIRED PACKAGES

```{r}
library(vegan); library(lubridate); library(dplyr); library(tidyr); library(nlme); library(ggplot2); library(sjPlot); library(ggpubr); library(fuzzyjoin); library(car); library(jcolors); library(glmmTMB)
```

### Reading in RDS objects

```{r}
df_prop_general <- readRDS("~/Desktop/CapuchinBehaviouralDiversity/df_prop_general_fruit.RDS")
df_prop_forage <- readRDS("~/Desktop/CapuchinBehaviouralDiversity/df_prop_forage_fruit.RDS")
df_combinedInd_general <- readRDS("~/Desktop/CapuchinBehaviouralDiversity/df_combinedInd_general_fruit.RDS")
df_combinedInd_forage <- readRDS("~/Desktop/CapuchinBehaviouralDiversity/df_combinedInd_forage_fruit.RDS")
df_census_filt <- readRDS("~/Desktop/CapuchinBehaviouralDiversity/df_census_filt.RDS") # this object is a df with census information and metadata for the entire study
```

### VISUALIZATIONS

#### GROUP SIZE OVER TIME

```{r}
groupTest <- df_census_filt %>%
             group_by(Group, CensusYear) %>%
             mutate(facet = case_when(Group %in% c("CP", "AD", "RM") ~ "CP to AD & RM",
                                      Group %in% c("GN", "BC", "GN2") ~ "GN to GN2 & BC",
                                      Group == "LV" ~ "LV",
                                      Group == "EX" ~ "EX")) %>%
            mutate(MeanGroupSize = mean(TotalGroupSize)) %>%
            ungroup()
            

groupSizeOverTime <- ggplot(data = groupTest, aes(x = factor(CensusYear), y = MeanGroupSize, color = Group, size = MeanGroupSize)) +
      geom_hline(yintercept = 10, linetype = "dashed", color = "grey") +
      geom_hline(yintercept = 20, linetype = "dashed", color = "grey") +
      geom_hline(yintercept = 30, linetype = "dashed", color = "grey") +
      geom_point() +
      scale_color_jcolors(palette = "pal7") +
      xlab("") + ylab("Mean Group Size\n") +
      ylim(1, 40) +
      theme_bw() +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title=element_text(size=18),
            axis.text=element_text(size=12)) +
     facet_grid(rows = vars(facet)); groupSizeOverTime

summary(df_census_filt$TotalGroupSize)
```  

### DATA ANALYSIS GROUP LEVEL GENERAL

```{r}
#visualizing raw data

rawDataGroupRichness <- ggplot(data = df_prop_general, aes(x = TotalGroupSize, y = S.obs)) +
  geom_jitter(size = 0.1) +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Observed Behavioral Richness\n") +
  xlim(0,40) + theme(axis.title=element_text(size=18), axis.text=element_text(size=12)); rawDataGroupRichness

rawDataGroupShannon <- ggplot(data = df_prop_general, aes(x = TotalGroupSize, y = shannon)) +
  geom_jitter(size = 0.1) +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Shannon Behavioral Diversity\n") +
  xlim(0,40) + theme(axis.title=element_text(size=18), axis.text=element_text(size=12)); rawDataGroupShannon
```

#### BEHAVIOURAL RICHNESS MODEL (Group Level)

```{r}
null_model <- glmmTMB(S.obs ~ 1 + (1 | Group/PropMonkeysInScan), data = df_prop_general, family = poisson)

model <- glmmTMB(S.obs ~ Temp_z + 
                   Rainfall_z + 
                   TotalGroupSize_z + 
                   PropFemales + 
                   biomass_monthly_kg_z +
                   (1 | Group/PropMonkeysInScan),
                   data = df_prop_general,
                   family = poisson)

AIC(null_model, model)

#           df      AIC
#null_model  3 32648.79
#model       8 32556.88

tab_model(model)
```

#### BEHAVIORAL SHANNON DIVERSITY MODEL (Group Level)

```{r}
null_model_shannon <- glmmTMB(shannon ~ 1 + (1 | Group/PropMonkeysInScan), data = df_prop_general, family = ziGamma, ziformula = ~ 1, start = list(beta = rep(0, 1)), control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS")))

model_shannon <- glmmTMB(shannon ~ Temp_z + 
                   Rainfall_z + 
                   TotalGroupSize_z + 
                   PropFemales + 
                   biomass_monthly_kg_z +
                 (1 | Group/PropMonkeysInScan),
                 data = df_prop_general,
                 family = ziGamma,
                 ziformula = ~ 1,
                 start = list(beta = rep(0, 6)),  # Adjust the length based on the number of predictors
                 control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS")))

AIC(null_model_shannon, model_shannon)

#                   df      AIC
#null_model_shannon  5 14340.47
#model_shannon      10 14237.89

tab_model(model, model_shannon, file = "~/Desktop/CapuchinBehaviouralDiversity/SuppTable2.doc")
```

We compared including proportion of monkeys in scan with number of monkeys in scan. First, in an AIC comparison, the model with Proportion of Monkeys in Scan has more explanatory power (delta AIC > 2). Second, in comparing the results of each model, the effect of our predictor of interest is nearly identical. Therefore, we proceeded with Proportion of Monkeys in Scan as our control for differences in sampling effort.

#### GROUP TESTING FOR AUTOCORRELATION

Autocorrelation of behaviours from one day to the next or one scan to the next can be an issue when interpreting p-values, as the non-independence of the data is not taken into account. We can check to see if this is a problem by looking at the autocorrelation in the residuals of the model.

```{r}
#First, fit we fit our linear mixed effects model using the glmmTMB function from the glmmTMB package.

mRich; m1

#Second, we extract the residuals from the fitted model using the resid() function.

mRich_resid <- resid(mRich)
mRich_resid_numeric <- unname(mRich_resid)

m1_resid <- resid(m1)
m1_resid_numeric <- unname(m1_resid)

#Third, create a time-related variable (e.g., the order of observations) and plot it

order_variable <- 1:nrow(df_prop_general)

plot(order_variable, mRich_resid_numeric, xlab = "Time", ylab = "Residuals", main = "Residuals Over Time")
plot(order_variable, m1_resid_numeric, xlab = "Time", ylab = "Residuals", main = "Residuals Over Time")

#Fourth, create an autocorrelation function (ACF) plot to assess the correlation between residuals at different lags
acf(mRich_resid_numeric, lag.max = 10)
acf(m1_resid_numeric, lag.max = 10)

# Assuming 'mRich_resid_numeric' is your numeric vector of residuals
n <- length(mRich_resid_numeric)
dw_statistic <- sum(diff(mRich_resid_numeric)^2) / sum(mRich_resid_numeric^2)

# Compare with critical values for significance
# For a two-sided test, compare with 1.5 (lower bound) and 2.5 (upper bound)
dw_significance <- dw_statistic < 1.5 | dw_statistic > 2.5

# Print results
cat("Durbin-Watson Statistic:", dw_statistic, "\n")
cat("Significant at 5% level?", dw_significance, "\n")

# Assuming 'm1_resid_numeric' is your numeric vector of residuals
n_m1 <- length(m1_resid_numeric)
dw_statistic_m1 <- sum(diff(m1_resid_numeric)^2) / sum(m1_resid_numeric^2)

# Compare with critical values for significance
# For a two-sided test, compare with 1.5 (lower bound) and 2.5 (upper bound)
dw_significance_m1 <- dw_statistic_m1 < 1.5 | dw_statistic_m1 > 2.5

# Print results
cat("Durbin-Watson Statistic:", dw_statistic_m1, "\n")
cat("Significant at 5% level?", dw_significance_m1, "\n")

# Results between 1.5 and 2.5 are considered evidence of slight autocorrelation. Our results are 1.8 for the richness model, and 1.6 for the Shannon model. These suggest slight positive autocorrelation of scans, but not at a level where we feel it is necessary to add a correlation structre to our models.
```

### PLOTTING MODEL PREDICTIONS (Group Level)

```{r}
richness_group_level_plot <- plot_model(model, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Behavioral Richness\n(Group Level)\n")
richness_group_level_plot

shannon_group_level_plot <- plot_model(model_shannon, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Behavioral Shannon Diversity\n(Group Level)\n")
shannon_group_level_plot

ggarrange(richness_group_level_plot, shannon_group_level_plot, labels = c("A", "B"))
```


### DATA ANALYSIS INDIVIDUAL LEVEL GENERAL 

```{r}
null_model_mRichInd_general <- glmmTMB(S.obs ~ 1 + (1 | TotalScans/MonkeyName), data = df_combinedInd_general, family = poisson)

df_combinedInd_general$Rainfall_z <- scale(df_combinedInd_general$Rainfall)
df_combinedInd_general$TempMax_z <- scale(df_combinedInd_general$TempMax)

mRichInd_general <- glmmTMB(S.obs ~ TempMax_z + 
                   Rainfall_z + 
                   TotalGroupSize_z + 
                   AgeClass + 
                   domLevel +
                   Sex +
                   biomass_monthly_kg_z +
                 (1 | TotalScans/MonkeyName),
                 data = df_combinedInd_general,
                 family = poisson)

AIC(null_model_mRichInd_general, mRichInd_general)


m1ind_null_general <-  glmmTMB(shannon ~ 1 +
                              (1 | TotalScans/MonkeyName), 
                              data = df_combinedInd_general,
                              family = ziGamma, ziformula = ~ 1)


m1ind_general <- glmmTMB(shannon ~ TempMax_z +
                              Rainfall_z +
                              AgeClass + 
                              TotalGroupSize_z +
                              Sex +
                              domLevel +
                              biomass_monthly_kg_z +
                              (1 | TotalScans/MonkeyName), 
                              data = df_combinedInd_general,
                              family = ziGamma, ziformula = ~ 1,
                              start = list(beta = rep(0, 12)),  # Adjust the length based on the number of predictors
                              control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS")))


AIC(m1ind_null_general, m1ind_general)

tab_model(mRichInd_general, m1ind_general, file = "~/Desktop/CapuchinBehaviouralDiversity/SuppTable3.doc")

ind_richness_level_plot <- plot_model(mRichInd_general, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("purple"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Behavioral Richness\n(Individual Level)\n"); ind_richness_level_plot

m1ind_general_shannon_level_plot <- plot_model(m1ind_general, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("purple"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Shannon Behavioral Diversity\n(Individual Level)\n"); m1ind_general_shannon_level_plot

ggarrange(ind_richness_level_plot, m1ind_general_shannon_level_plot, labels = c("A", "B"))
```

### Does social aggression increase with group size?

```{r}
df_prop_general #general ethogram

ggplot(data = df_prop_general, aes(x = TotalGroupSize, y = SAG)) +
  geom_jitter() +
  theme_minimal()

mSAG_null <- glmmTMB(SAG ~ 1 +
                   (1 | Group/PropMonkeysInScan),
                   data = df_prop_general,
                   family = poisson)

mSAG <- glmmTMB(SAG ~ Temp_z + 
                   Rainfall_z + 
                   TotalGroupSize_z + 
                   PropFemales + 
                   biomass_monthly_kg_z +
                   (1 | Group/PropMonkeysInScan),
                   data = df_prop_general,
                   family = poisson)

AIC(mSAG_null, mSAG)

#          df      AIC
#mSAG_null  3 5629.661
#mSAG       8 5562.256

df_combinedInd_general$Rainfall_z <- scale(df_combinedInd_general$Rainfall)
df_combinedInd_general$Temp_z <- scale(df_combinedInd_general$TempMax)
df_combinedInd_general$domLevel <- factor(df_combinedInd_general$domLevel, levels = c("Unassessed", "low", "mid", "high"))

mSAG_ind_null <- glmmTMB(SAG ~ 1 +
                   (1 | TotalScans/MonkeyName),
                   data = df_combinedInd_general,
                   family = poisson)

mSAGInd_general <- glmmTMB(SAG ~ Temp_z + 
                   Rainfall_z + 
                   AgeClass + 
                   TotalGroupSize_z +
                   TotalGroupSize_z*domLevel +
                   domLevel +
                   biomass_monthly_kg +
                   Sex +
                   (1 | TotalScans/MonkeyName),
                   data = df_combinedInd_general,
                   family = poisson)

AIC(mSAG_ind_null, mSAGInd_general)

#                df      AIC
#mSAG_ind_null    3 5317.406
#mSAGInd_general 17 5043.079

tab_model(mSAG, mSAGInd_general, file = "~/Desktop/CapuchinBehaviouralDiversity/SuppTable6.doc")

SAG_group_level_plot <- plot_model(mSAG, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Number of Monkeys\nExhibitng Agonistic Behaviors\n(Group Level)\n"); SAG_group_level_plot

SAG_ind_level_plot <- plot_model(mSAGInd_general, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Agonism Scans at Individual Level\n"); SAG_ind_level_plot

SAG_ind_level_plot_dom <- plot_model(mSAGInd_general, 
                                 type = "pred", 
                                 terms = c("TotalGroupSize_z", "domLevel"),
                                 title = "",
                                 colors = c("red", "cornflowerblue", "darkorange", "forestgreen"),  # Colors for different levels of domLevel
                                 ci.lvl = 0.95,  # Confidence interval level
                                 alpha = 0.3) +  # Transparency of CI intervals
                                 theme_minimal() +
                                 labs(x = "\nTotal Group Size (z-scored)", 
                                      y = "Predicted Agonism Scans at Individual Level\n"); SAG_ind_level_plot_dom



ggarrange(SAG_group_level_plot, SAG_ind_level_plot, labels = c("A", "B"))
```

### DATA ANALYSIS GROUP LEVEL - FORAGE

```{r}
rawDataGroupRichnessForaging <- ggplot(data = df_prop_forage, aes(x = TotalGroupSize, y = S.obs)) +
  geom_jitter() +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Observed Behavioural Richness\n"); rawDataGroupRichnessForaging

rawDataGroupShannonForaging <- ggplot(data = df_prop_forage, aes(x = TotalGroupSize, y = shannon)) +
  geom_jitter() +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Shannon Behavioural Diversity\n"); rawDataGroupShannonForaging

ggarrange(rawDataGroupRichnessForaging, rawDataGroupShannonForaging)

#null model
null_mRich_forage <- glmmTMB(S.obs ~ 1 + 
                              (1 | Group/PropMonkeysInScan),
                              data = df_prop_forage,
                              family = poisson)

#model 
mRich_forage <- glmmTMB(S.obs ~ Temp_z +
                              Rainfall_z +
                              TotalGroupSize_z +
                              PropFemales +
                              biomass_monthly_kg + 
                              (1 | Group/PropMonkeysInScan),
                              data = df_prop_forage,
                              family = poisson)

#compare null to model
AIC(null_mRich_forage, mRich_forage)

#                  df      AIC
#null_mRich_forage  3 27458.42
#mRich_forage       8 27450.78

tab_model(mRich_forage, file = "~/Desktop/CapuchinBehaviouralDiversity/SuppTable4.doc")

null_m1_forage <-  glmmTMB(shannon ~ 1 +
                              (1 | Group/PropMonkeysInScan),
                              data = df_prop_forage,
                              family = ziGamma,
                            ziformula = ~ 1, start = list(beta = rep(0, 1)))
                           
m1_forage <-  glmmTMB(shannon ~ #Temp_z +
                              #Rainfall_z +
                              TotalGroupSize_z +
                              #PropFemales +
                              biomass_monthly_kg +
                              (1 | Group/PropMonkeysInScan),
                              data = df_prop_forage,
                              family = ziGamma,
                              ziformula = ~ 1, start = list(beta = rep(0, 3)), control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS")))

AIC(null_m1_forage, m1_forage)

#                  df      AIC
#null_mRich_forage  3 27458.42
#mRich_forage       8 27450.78

mRich_forage_plot <- plot_model(mRich_forage, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Foraging Behavioral Richness\n(Group Level)\n"); mRich_forage_plot


ggarrange(mRich_forage_plot, SAG_group_level_plot, labels = c("A", "B"))
```

### DATA ANALYSIS INDIVIDUAL LEVEL FORAGING 

```{r}

rawDataIndRichnessForaging <- ggplot(data = df_combinedInd_forage, aes(x = TotalGroupSize, y = S.obs)) +
  geom_jitter() +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Observed Behavioural Richness (Foraging)\n"); rawDataIndRichnessForaging

rawDataIndShannonForaging <- ggplot(data = df_combinedInd_forage, aes(x = TotalGroupSize, y = shannon)) +
  geom_jitter() +
  theme_minimal() +
  xlab("\nGroup Size") + ylab("Shannon Behavioural Diversity (Foraging)\n"); rawDataIndShannonForaging

ggarrange(rawDataIndRichnessForaging, rawDataIndShannonForaging)

null_model_mRichInd_forage <- glmmTMB(S.obs ~ 1 + 
                            (1 | TotalScans/MonkeyName),
                       data = df_combinedInd_forage,
                       family = poisson)

df_combinedInd_forage$Rainfall_z <- scale(df_combinedInd_forage$Rainfall)
df_combinedInd_forage$TempMax_z <- scale(df_combinedInd_forage$TempMax)

mRichInd_forage <- glmmTMB(S.obs ~ TempMax_z +
                              Rainfall_z+
                              AgeClass + 
                              TotalGroupSize_z +
                              domLevel +
                              Sex +
                              TotalGroupSize_z*domLevel + (1 | TotalScans/MonkeyName),
                              data = df_combinedInd_forage,
                              family = poisson)

AIC(null_model_mRichInd_forage, mRichInd_forage)

m1ind_null_forage <-  glmmTMB(shannon ~ 1 +
                              (1 | TotalScans/MonkeyName), 
                              data = df_combinedInd_forage,
                              family = ziGamma, ziformula = ~ 1)

m1ind_forage <- glmmTMB(shannon ~ TempMax_z +
                              Rainfall_z +
                              AgeClass + 
                              TotalGroupSize_z +
                              Sex +
                              domLevel +
                              TotalGroupSize_z*domLevel +
                              (1 | TotalScans/MonkeyName), 
                              data = df_combinedInd_forage,
                              family = ziGamma, ziformula = ~ 1,
                              start = list(beta = rep(0, 14)),  # Adjust the length based on the number of predictors
                              control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS")))

AIC(m1ind_null_forage, m1ind_forage)
tab_model(m1ind_forage)

tab_model(mRichInd_forage, m1ind_forage, file = "~/Desktop/CapuchinBehaviouralDiversity/SuppTable5.doc")

mRich_forage_plot <- plot_model(mRich_forage, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Foraging Behavioral Richness\n(Group Level)\n"); mRich_forage_plot

mRich_Id_forage_plot <- plot_model(m1ind_forage, 
                                        type = "pred", 
                                        terms = c("TotalGroupSize_z"),
                                        title = "",
                                        colors = c("cornflowerblue"),  # Line color and CI color
                                        ci.lvl = 0.95,  # Confidence interval level
                                        alpha = 0.3) +  # Transparency of CI intervals
                                        theme_minimal() +
                                        labs(x = "\nTotal Group Size (z-scored)", 
                                             y = "Predicted Shannon Foraging Diversity\n(Individual Level)\n"); mRich_Id_forage_plot

ggarrange(mRich_forage_plot, mRich_Id_forage_plot, labels = c("A", "B"))
```
